---
title: "B2"
author: "EMT"
date: "2023-10-02"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## B2 Homework

#### Evidence of Data Fabrication

```{r echo = FALSE, warning = FALSE, message = FALSE}
setwd("~/myrepo GH/B HOMEWORK/B2")
library(tidyverse)
gay <- read_csv("gay.csv")
head(gay)
dim(gay)
ccap <- read.csv("CCAP.csv")
head(ccap)
dim(ccap)
```

**Question 1**

```{r}
dens_gay <- ggplot(data = gay)
dens_gay <- dens_gay + geom_histogram(mapping = aes(y=..density.., x = feel1 ), binwidth=10)
dens_gay <- dens_gay + theme_classic() + xlab("Feelings towards gay people, gay") + ylab("Density")         
dens_gay

dens_ccap <- ggplot(data = ccap)
dens_ccap <- dens_ccap + geom_histogram(mapping = aes(y=..density.., x = feel ), binwidth=10)
dens_ccap <- dens_ccap + theme_classic() + xlab("Feelings towards gay people, CCAP") + ylab("Density")         
dens_ccap



```

Yes, the distribution of the data from the gay marriage experiment is almost identical to the distribution of the CCAP data.

**Question 2**

```{r}
gay %>%
  summarise(
            mean_feel1 = mean(feel1, na.rm = TRUE),
            median_feel1 = median(feel1, na.rm = TRUE),
            sd_feel1 = sd(feel1, na.rm = TRUE)
            )
ccap %>%
  summarise(
            mean_feel = mean(feel, na.rm = TRUE),
            median_feel = median(feel, na.rm = T),
            sd_feel = sd(feel, na.rm = TRUE),
  )
```

Yes, the distributions of the two datasets are practically identical.

**Question 3**

I would expect that the baseline attitudes towards same-sex marriage would be more positive than the national average, such that the distribution was shifted more towards the right (towards 100) seeing as Los Angeles county is a more liberal area of the United States and thus not representative of the whole American population.

**Question 4**

```{r}

feel1v2 <- ggplot(data=gay[gay$treatment=="No Contact",])
feel1v2 <- feel1v2 + geom_point(mapping=aes(x=feel1, y=feel2))
feel1v2

feel1v3 <- ggplot(data=gay[gay$treatment=="No Contact",])
feel1v3 <- feel1v3 + geom_point(mapping=aes(x=feel1, y=feel3))
feel1v3

feel1v4 <- ggplot(data=gay[gay$treatment=="No Contact",])
feel1v4 <- feel1v4 + geom_point(mapping=aes(x=feel1, y=feel4))
feel1v4

```

Yes, the respondents' answers are very similar over time. The error-minimizing line would be the 45 degree line, in the data it does look like the observations follow this line, although there is some random noise around it. This could suggest, as the assignment writes, that the researchers have reused the feel variable from the CCAP data and added random noise to create the "control group" for each round of feel-survey.

**Question 5**

```{r}
gay %>%
  filter(treatment=="No Contact") %>%
summarise(
  cor2 = cor(feel1,feel2, use="complete.obs"),
  cor3 = cor(feel1,feel3, use="complete.obs"),
  cor4 = cor(feel1,feel4, use="complete.obs")

)
```

The answers for the control group of feel1 are highly correlated with the answers to feel2, feel3 and feel4 with correlation coefficients above .95 in each case.

**Question 6**

Yes, there is evidence suggesting that the data used in the gay marriage experiment is fabricated.

### Predicting Course Grades Using Midterm Scores Part 1: Predicting Final Exam Scores

```{r}
grades <- read.csv("grades.csv")
head(grades)
dim(grades)
```

**Question 1**

a) Each observation in the dataset represents a student, the variables reflect the specific student's performance in their class.

b) We will use the midterm test scores, i.e. `midterm` as our X variable. It is a non-binary variable as it can take values between 0 and 100.

c) As stated in the assignment title, we want to predict final exam scores and thus our Y variable will be `final`. This variable is non-binary as it can values between 0 and 100.

d) The correlation coefficient between the midterm test scores and the finals test scores is positive and quite strong at .716.

```{r}
cor_xy <- cor(grades$midterm,grades$final)
cor_xy
```

**Question 2**

a) Calculating the fitted line

```{r}
fit <- lm(final <- final ~ midterm, data = grades)
fit
summary(fit)
```

b) The fitted line is `final = -6.01 + .97*midterm`

c) The relationship can be visualized as follows:

```{r}
testscores <- ggplot(data=grades,mapping=aes(x=midterm, y=final) )
testscores <- testscores + geom_point() 
testscores <- testscores + geom_smooth(method="lm", se=FALSE, colour="red")
testscores <- testscores + xlab("Midterm Scores") + ylab("Finals Scores")
testscores
```

**Question 3**

a) My best guess would be using the conditional expectation, that is, conditioning on my midterm score of 80 points, my finals score would be 71.63 points.

b) If my midterm score is 90 points, then my final exam score will be 81.33 points.

c) An increase of 10 midterm points would result in an increase of 10\*0.9704=9.704 points on the final exam.

```{r}
y_80 <- -6.0059+0.9704*80
y_80 

y_90 <- -6.0059+0.9704*90
y_90 

delta_y_10 <- 10*.9704
delta_y_10
```

**Question 4**

```{r}
R2 <- cor_xy^2
R2
```

The R-squared is .51 and can be interpreted as the share of the variation in finals test scores that is explained by variation in mid term test scores. It reflects how much better the linear regression model is at predicting finals test scores compared to taking the sample mean. Just above half of the variation in finals test scores can be explained by variation in the midterm scores. The R-squared can also be retrieved from the summarized output of the linear regression.
