---
title: "B1"
author: "Emma Terreni"
date: "2023-09-19"
output: 
  html_document: 
    keep_md: true
---

## B1 Homework

#### Indiscriminate Violence and Insurgency

```{r echo = FALSE, warning = FALSE, message = FALSE}
setwd("~/myrepo GH/B HOMEWORK/B1")
library(tidyverse)
chechen <- read_csv("chechen.csv")
head(chechen)
dim(chechen)

chechen
```

**Question 1**

```{r}
    chechen %>%
        count(fire)

```

159 villages in the data were attacked by Russians and an equivalent 159 villages were not attacked by Russians.

**Question 2**

```{r}
chechen %>%
  filter(fire == 1) %>%
  group_by(groznyy) %>%
  summarise(
            meandeaths = mean(deaths, na.rm = TRUE),
            mediandeaths = median(deaths, na.rm = TRUE),
            n = n()
            )
 
```

When only looking at the villages where there were Russian attacks, it is apparent that the villages in Groznyy experienced a greater number of deaths compared to the villages outside of Groznyy, bot when looking at the mean ( a difference of 2.2 deaths on average) and the median (a difference of 3 deaths). It is important to note, that there were only 7 villages in Groznyy exposed to Russian attacks compared to the 152 outside of Groznyy.

**Question 3**

```{r}
chechen %>%
  group_by(fire) %>%
  summarise(
            meanpostattacks = mean(postattack, na.rm = TRUE),
            postattack_q25 = quantile(x = postattack, probs = 0.25, na.rm = T),
            postattack_q50 = median(postattack, na.rm = TRUE),
            postattack_q75 = quantile(x = postattack, probs = 0.75, na.rm = T),
            n = n()
            )
```

The average number of insurgent attacks after Russian fire for villages hit by artillery fire is 1.5 while the corresponding number is 2.05 for the villages not hit.

The first and second quartile are identical for villages hit and villages not hit, while the 3rd quartile is one death higher for the villages that were not hit by Russian fire. This suggests that the villages most severely affected, that were not hit, have higher numbers of deaths than the most severely hit villages, that were not affected by Russian artillery fire. That might also be what is reflected in the difference in means between the two groups.

This would suggest that the fires dampened the insurgent attacks, but since we have not compared to the number of attacks before fire, we cannot say anything causal about the descriptive statistics. It might just be, that the villages not hit by Russian fire experience more attacks on average than the villages hit. The table above says nothing about the change in insurgent attacks. I would not conclude that indiscriminate violence reduces insurgent attacks yet.

**Question 4**

```{r}
chechen %>%
  group_by(fire) %>%
  summarise(
            meanpreattacks = mean(preattack, na.rm = TRUE),
            preattack_q25 = quantile(x = preattack, probs = 0.25, na.rm = T),
            preattack_q50 = median(preattack, na.rm = TRUE),
            preattack_q75 = quantile(x = preattack, probs = 0.75, na.rm = T),
            n = n()
            )
```

The average number of attacks before Russian fire was almost identical for villages hit by Russian fire and for villages not hit. This suggests that, as far as frequency of insurgent attacks, the two groups were similar before Russian artillery fire. Lyall (2009) states that the Russian artillery strikes are uncorrelated with key spatial and demographic characteristics, making it a sound instrument. If we believe that the Russian artillery fires were carried out randomly and that the villages experiencing Russian fire are comparable to the villages that didn't, we are able to draw causal conclusions. The drop in insurgent attacks in the villages affected by Russian artillery fire is, on average, notably larger than the drop in villages not affected.

**Question 5**

Among the villages shelled by Russians, the number of insurgent attacks did not increase after the artillery attack. In fact, the number of insurgent attacks decreased on average. When comparing to the average decrease in attacks in non-shelled villages, the decrease in the shelled villages is substantial. When looking at the distribution of the differences in attacks, we see that the difference in means is mainly driven by a few villages experiencing a vast drop in attacks after Russian fire, as well as a generally more skewed distribution compared to the villages not shelled.

```{r}
chechen %>%
  mutate(diffattacks = postattack - preattack) %>%
  group_by(fire) %>%
  summarise(meandiffattacks = mean(diffattacks, na.rm = TRUE),
            mediandiffattacks = median(diffattacks, na.rm = TRUE)
            )

```

```{r}
chechen_new <- chechen %>%
  mutate(diffattacks = postattack - preattack) 

head(chechen_new)

hist2 <- ggplot(
  data=chechen_new, 
  mapping = aes(
    x = diffattacks,
    group = as.factor(fire),
    fill = as.factor(fire)
    ) 
  )

hist2 + geom_histogram(
  position = "dodge",
  binwidth = 0.5
  )
```

**Question 6**

The mean difference in the `difattacks` variable between villages shelled and villages not shelled is 0.52 attacks, meaning that there were, on average .52 fewer insurgent attacks in the villages shelled than the villages not shelled after the Russian fire. The difference-in-differences estimator has shown that the change in insurgent attacks in shelled villages is greater in magnitude than the change in insurgent attacks in villages not shelved. Accepting the assumption that villages shelled and not shelled do not vary systematically and that the Russian artillery fire was carried out randomly, i.e. not correlated with any characteristics specific to the villages, this analysis supports the claim that indiscriminate violence reduces insurgency attacks.

The validity of this analysis is improved compared to the previous questions since this analysis takes the change into account and not just the post Russian fire data. Using information on insurgency attacks before the Russian fire allows us to compute whether the Russian fires actually made a difference to the number of insurgency attacks or not. We needed pre-treatment data, so to speak, to evaluate any type of effect before even beginning to talk about causality.

```{r}
mean(chechen_new$diffattacks[chechen_new$fire == 1])-mean(chechen_new$diffattacks[chechen_new$fire == 0])
```

### Inequality of Success in Online Music Markets

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(readr)
users1 <- read_csv("users1.csv")
users2 <- read_csv("users2.csv")
songs1 <- read_csv("songs1.csv")
songs2 <- read_csv("songs2.csv")

```

**Question 1**

Proportion of users assigned separately for the Social Influence and the Independent condition.

```{r}
users1_new <- users1 %>%
  mutate(expnr = 1)

users2_new <- users2 %>%
  mutate(expnr = 2)

users_full <- rbind(users1_new,users2_new)

view(users_full)

  table1 <- prop.table(table(users_full$expnr, users_full$world_id), margin = 1)
  round(table1,3)
  
  tabledata1 <- prop.table(table(users1$world_id))
  tabledata2 <- prop.table(table(users2$world_id))

  tabledata1
  tabledata2
```

The proportions within each experiment are practically identical, with 33% assigned to the Social Influence condition and 67% to the Independence condition in Experiment 1. The equivalent proportions are 32% and 68% for experiment 2.

**Question 2**

The average number of downloads (assuming this is equivalent to purchasing the song) per user.

```{r}

usersfull_temp <- users_full %>%
  group_by(expnr, world_id) %>%
  summarise(
    mean_download = mean(download)
    )

usersfull_temp


```

The average number of downloads per user in Experiment 1 is 42.2 for the people assigned the social influence condition and 47.15 for the people assigned the Independence condition.

The average number of downloads per user in Experiment 2 is 301.2 for the people assigned the social influence condition and 124.1 for the people assigned the independence condition.

```{r}
# calculating number of workers
usersnr <- users_full %>%
  group_by(expnr, world_id) %>%
  summarise(n_users = n())
head(usersnr)

# calculating total number of times a song was listened to
songs1_new <- songs1 %>%
  mutate(expnr = 1)
songs2_new <-songs2 %>%
  mutate(expnr = 2)
songs_full <- rbind(songs1_new,songs2_new)

songsnrtemp <- songs_full %>%
  group_by(expnr) %>%
  summarise(
            listen_soctotal = sum(listen_soc),
            listen_indeptotal = sum(listen_indep)
            )
songsnrtemp

songsnr <- songsnrtemp %>%
    pivot_longer(
    cols = starts_with("listen"), 
    names_to = "world_id", 
    values_to = "listen_total"
    )

songsnr$world_id[which(songsnr$world_id=="listen_soctotal")] <- "1"
songsnr$world_id[which(songsnr$world_id=="listen_indeptotal")] <- "9"
songsnr$world_id <- as.numeric(songsnr$world_id)

# merging users and song data
totals <- merge(usersnr, songsnr, by=c("expnr","world_id"))
totals <- totals %>%
  mutate(listen_per_user = listen_total/n_users)
totals

```

The average listens per user in Experiment 1 is 3.51 for the people assigned the social influence condition and 3.74 for the people assigned the Independence condition.

The average listens per user in Experiment 2 is 3.56 for the people assigned the social influence condition and 3.9 for the people assigned the independence condition.

It seems odd that there are less listens than downloads per user. If we use the songs-dataset for downloads as well as listens, there will be lower downloads than listens, leading me to believe that the relevant downloads-data is from the songs-data and not the users-data.

**Question 3**

```{r}
library(ineq)

songs_full %>%
  group_by(expnr) %>%
  summarise(
    gini_down_soc = ineq(down_soc),
    gini_down_indep = ineq(down_indep),
    gini_listen_soc = ineq(listen_soc),
    gini_listen_indep = ineq(listen_indep)
    )
```

The results support the hypothesis that inequality in both listens and downloads is higher under the social influence condition than the independent condition. Furthermore, in Experiment 2 where the songs were ranked according to popularity as well as displaying number of downloads for each song, the inequality is substantially larger.

**Question 4**

```{r}
users_full %>%
  group_by(expnr, world_id) %>%
  summarise(
    mean_web = mean(web, na.rm = TRUE),
    mean_visit = mean(visit, na.rm = TRUE),
    mean_purchase = mean(purchase, na.rm = TRUE),
    n = n()
  )
```

The groups are on average practically identical across experiment and condition type. Additionally, the group sizes for each condition are similar.

**Question 5**

```{r}
exp1effect_d <- ineq(songs1$down_soc)-ineq(songs1$down_indep)
exp2effect_d <- ineq(songs2$down_soc)-ineq(songs2$down_indep)

exp1effect_d
exp2effect_d

exp1effect_d - exp2effect_d

```

```{r}
exp1effect_l <- ineq(songs1$listen_soc)-ineq(songs1$listen_indep)
exp2effect_l <- ineq(songs2$listen_soc)-ineq(songs2$listen_indep)

exp1effect_l
exp2effect_l

exp1effect_l - exp2effect_l

```

The difference in the change in Gini-coefficient is .18 and .28 gini-points for downloads and listens respectively. The between study comparison reflects the extra increase in inequality from exposing the users to an ordering of the songs by popularity. One could suspect this between study to have less internal validity simply from the fact that we cannot be sure whether the experiments were conducted similarly and thus are believably comparable. However, question 4 revealed that the groups in the two experiments look similar in measurable characteristics which strengthens the validity of the between study comparison.
