---
title: "B3"
author: "EMT"
date: "2023-10-16"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## B3 HOMEWORK

### Chapter 1 - Problem 22

The probability of the eldest 3 children being the 3 girls is 5%.

```{r}
sims <- 100000
children <- c("girl","girl","girl","boy","boy","boy")
event <- 0


for (i in 1:sims) {
  orders <- sample(x=children, size=6, replace=FALSE)
  if (orders[1]=="girl" & orders[2]=="girl" & orders[3]=="girl"){
    event <- event+1
  }
}

answer <- event/sims
answer
```

### Chapter 1 - Problem 34

The probability of getting 5 of each dice-value in a throw of 30 dice is 0,04%

```{r}
sims <- 100000
dice <- c(1,2,3,4,5,6)

rollcount <- 0
favoutcome <- 0


for (i in 1:sims) {
      roll <- sample(x=dice, size=30, replace=TRUE)
      for(v in 1:6){
      a <- sum(roll==v)
      if(a==5){
        rollcount <- rollcount + 1
      }
      } 
      if(rollcount==6){
        favoutcome <- favoutcome + 1 
      }
      rollcount <-0
     }

answer <- favoutcome/sims
answer

```

### Chapter 2 - Problem 2

The probability that the woman pregnant with twin boys is carrying identical twins is 50%.

```{r}

twintype <- c("frat","frat","ident")
fratcombo <- c("BB","BG","GB","GG")
identcombo <- c("BB","GG")

identBBcount <- 0
BBcount <- 0

for (i in 1:sims){
  twintypei <- sample(twintype, 1)
  if(twintypei=="frat"){
    sexcombo <- sample(fratcombo, 1)
  }
  else{
    sexcombo <- sample(identcombo, 1)
  }
  if(twintypei=="ident" & sexcombo == "BB"){
    identBBcount <- identBBcount + 1
  }
  if(sexcombo == "BB"){
    BBcount <- BBcount+1
  }
}

answer <- identBBcount/BBcount
answer
```

### Chapter 2 - Problem 32

For all three scenarios listed, the probability is 2/3 = 66,6%.

The event A\>B is independent of B\>C and should thus have the same probability whether conditioning on B\>C or not.

The event B\>C is not independent of C\>D and the probability should be lower for B\>C given that C\>D.

```{r}
a <- c(4,4,4,4,0,0)
b <- c(3,3,3,3,3,3)
c <- c(6,6,2,2,2,2)
d <- c(5,5,5,1,1,1)

acount <- 0
bcount <- 0
ccount <- 0
dcount <- 0

aacount <- 0
abcount <- 0

for (i in 1:sims){
  aroll <- sample(x=a, size=1, replace = TRUE)
  broll <- sample(x=b, size=1, replace = TRUE)
  croll <- sample(x=c, size=1, replace = TRUE)
  droll <- sample(x=d, size=1, replace = TRUE)
  
  # PART A
  if(aroll>broll){
    acount <- acount + 1
  }
   if(broll>croll){
    bcount <- bcount + 1
   }
   if(croll>droll){
    ccount <- ccount + 1
   }
   if(droll>aroll){
    dcount <- dcount + 1
   }
 
   # PART B
  if(broll>croll){
    if(aroll>broll){
      aacount <- aacount + 1
      }
    }
  if(broll>croll){
    if(croll>droll){
      abcount <- abcount + 1
      }
    }
}

aanswer <- acount/sims
banswer <- bcount/sims
canswer <- ccount/sims
danswer <- dcount/sims
aanswer
banswer
canswer
danswer

baanswer <- aacount/bcount
baanswer

bbanswer <- abcount/ccount
bbanswer

```

### Chapter 2 - Problem 38

There is a 1/7 = 14,3% probability of getting the car from not switching and a 2/7=28,3% probability of getting the car when switching.

```{r}
sims <- 100000
doors <- c("goat", "goat", "goat", "goat", "car", "goat", "goat")
result.switch <- result.noswitch <- rep(NA, sims)

  for (i in 1:sims) {
  ## randomly choose the initial door
  initialdoor <- sample(1:7, size = 1)
  result.noswitch[i] <- doors[initialdoor]
  remain <- doors[-initialdoor] # remaining six doors after initial door is chosen
  ## Monty chooses three doors with a goat
  if (doors[initialdoor] == "car"){ # initial door holds the car
    monty <- sample(1:6, size=3)
  } else{ # 5 goats and one car left
    car_placement <- which(grepl("car", remain))
    tmp <- c(1,2,3,4,5,6)
    monty <- sample(tmp[tmp!=car_placement], 3)
  }
    remain_stp2 <- remain[-monty] 
    result.switch[i] <- sample(remain_stp2, 1)
  }
  
mean(result.noswitch == "car") 
mean(result.switch == "car")
```
