---
title: 'A1: POLI210'
author: "EMT"
date: "2023-09-05"
output: 
  html_document: 
    keep_md: true
---
  
  Exercises 4.4 of "R for Data Science"
  09-06-2023
  
  
  Exercise 1
  Why does the code not work?
  
  my_variable <- 10
  my_varıable
  
    ```{r, eval=TRUE}
    
  my_variable <- 10
  my_variable
  
    ```

  The code doesn't work because the i in "my_variable" is replaced by a different
  letter (no dot), and so R cannot recognize the variable, the code is referring to.
  
  Exercise 2
  Tweak the following R commands so that they run correctly.
  The correct running code has been inserted. 
  

  install.packages("tidyverse")
    ```{r, eval=TRUE}
    
  library(tidyverse)
    
    ```
 
  Tweak: Install the package first
  
    ```{r, eval=TRUE}
    
  ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))
    
    ```
    
  Tweak: dota changed to data
  
    ```{r, eval=TRUE}
    
  filter(mpg, cyl == 8)
    
    ```
  
  Tweak: The command filter demands "==" syntax + spell filter correctly
  
    ```{r, eval=TRUE}
    
  filter(diamonds, carat > 3)
    
    ```
    
    Tweak: The dataset is called diamonds, not diamond - now specified correctly
  
  Exercise 3
  Alt + Shift + K opens an overview of shortcuts.
  I can get to the same place by going to Tools -> "Keyboard Shortcuts Help"
  
  